Classless Java

### 2016-03-22 OOPSLA16/sections/FutureWork.tex
Some reorganization, and some sad sacrifice to save some space

\paragraph{A real IB language}
In this article we presented the concept of IB as a programming pattern over Java.
We do not wish to define here a complete IB language, just to show the reader IB as
a promising way to design a new oo language.
The first task is to define the set of abstract state operations.
Of course we need getters; \Q@withX@ methods seams to be pretty convenient, and they enable a
clean function programming style that would be otherwise too cumbersome.

Currently we are offering both setters and Fluent setters.
We believe be fluent setters to be sufficient; we could drop the concept of (non fluent)setters.

Manually defining abstract state methods can be verbose, we could offer a convenient syntax to 
declare get/set/with.
For example
\begin{lstlisting}
obj interface Foo{
  String bar;//== String bar(); Foo withBar(String val);
  var String beer;//also fluent setter
  }
\end{lstlisting}

We may also want to offer a compact syntax for property updaters and generalized with methods;
\begin{lstlisting}
obj interface Foo{
  updater Bar, Beer//== Foo set(Bar val); Foo set(Beer val);
  with Foo, Beer//== Foo with(Bar val); Foo with(Beer val);
  }
\end{lstlisting}
Finally, certain methods may have a specific meaning, for example
if a private method called \Q@init@ is present, it will be called after construction,
and it may be handled specially during multiple implementations, so that all relevant \Q@init@ methods are
called in some order.

>>>>

%\paragraph{A real IB language}
%In this article we presented the concept of IB as a programming pattern over Java.
%We do not wish to define here a complete IB language, just to show the reader IB as
%a promising way to design a new oo language.
%The first task is to define the set of abstract state operations.
%Of course we need getters; \Q@withX@ methods seams to be pretty convenient, and they enable a
%clean function programming style that would be otherwise too cumbersome.
%
%Currently we are offering both setters and Fluent setters.
%We believe be fluent setters to be sufficient; we could drop the concept of (non fluent)setters.
%
%Manually defining abstract state methods can be verbose, we could offer a convenient syntax to 
%declare get/set/with.
%For example
%\begin{lstlisting}
%obj interface Foo{
%  String bar;//== String bar(); Foo withBar(String val);
%  var String beer;//also fluent setter
%  }
%\end{lstlisting}
%
%We may also want to offer a compact syntax for property updaters and generalized with methods;
%\begin{lstlisting}
%obj interface Foo{
%  updater Bar, Beer//== Foo set(Bar val); Foo set(Beer val);
%  with Foo, Beer//== Foo with(Bar val); Foo with(Beer val);
%  }
%\end{lstlisting}
%Finally, certain methods may have a specific meaning, for example
%if a private method called \Q@init@ is present, it will be called after construction,
%and it may be handled specially during multiple implementations, so that all relevant \Q@init@ methods are
%called in some order.

### 2016-03-22 OOPSLA16/sections/Introduction.tex
Some reorganization, and some sad sacrifice to save some space

\marcoT{
A key challenge lies in how to model state, which is fundamental to
have stateful objects. 
Programmers do not define objects directly, but
delegate the task to \emph{object interfaces}, whose role is similar
to non abstract classes in CB languages.
All abstract operations in an object interface will be interpreted as \emph{abstract state operations}. The
abstract state operations include various common utility methods (such
as getters and setters, or clone-like methods). In the presence of
subtyping, such operations often require special care, as their types
need to be refined. Object interfaces provide support for
type-refinement and can automatically produce code that deals with
type-refinement adequately, while CB (as in Java, Scala and C\#) require 
verbose explicit type-refinement.
We believe that such verbosity hindered and slowed down
the discovery of useful programming pattens involving type-refinement\cite{ trivial solution to the EP}.
}

>>>>

\marcoT{
A key challenge lies in how to model state, which is fundamental to
have stateful objects. 
Programmers do not define objects directly, but
delegate the task to \emph{object interfaces}, whose role is similar
to non abstract classes in CB languages.
All abstract operations in an object interface will be interpreted as \emph{abstract state operations}. The
abstract state operations include various common utility methods (such
as getters and setters, or clone-like methods). In the presence of
subtyping, such operations often require special care, as their types
need to be refined. Object interfaces provide support for
type-refinement and can automatically produce code that deals with
type-refinement adequately, while CB (as in Java, Scala and C\#) require 
verbose explicit type-refinement.
We believe that such verbosity hindered and slowed down
the discovery of useful programming pattens involving type-refinement\cite{trivialsolutiontotheEP}.
}

### 2016-03-22 OOPSLA16/sections/RelatedWork.tex
Some reorganization, and some sad sacrifice to save some space

\begin{comment}
\subsection{Type-Directed Translations/Syntactic Sugar}
\marco{I'm tring to merge this and the next one}
Language extensions are often implemented as syntactic sugar of the base
language. For example, Scala compiler supports XML syntax in normal Scala code
directly (after Scala ?, users need to import \texttt{scala-xml} library
manually). However, this approach is hard in terms of implementation, because it
requires extending the compiler. Also, this approach does not support combining
multiple extensions into one.

>>>>

\begin{comment}

\paragraph{Comparison with other Lombok annotations}\marco{if we keep it, will go in related work}
The Lombok project provides a set of predefined annotations, including constructor
generators similar as ours (e.g., \Q:@NoArgsConstructor:,
\Q:@RequiredArgsConstructor: and \Q:@AllArgsConstructor:). They
generate various kinds of constructors for \emph{classes}, with or without
constructor arguments. This set of annotations is of great use, especially when
used together with other features provided in Lombok (e.g.,
\Q:@Data:). Moreover, the implementation of these annotations in Lombok
gives us hints on how to implement \mixin. However, none of these annotations
can model what we are doing with \mixin - generating constructor-methods
(\textbf{of}) for \emph{interfaces}. Apart from constructors, \mixin also
provides other convenient features (including generating fluent setters, type
refinement, etc), which the base Lombok project does not provide.
Finally, while \mixin is formalized, none of Lombok's annotations have been
studied in a formal way.

\subsection{Type-Directed Translations/Syntactic Sugar}
\marco{I'm tring to merge this and the next one}
Language extensions are often implemented as syntactic sugar of the base
language. For example, Scala compiler supports XML syntax in normal Scala code
directly (after Scala ?, users need to import \texttt{scala-xml} library
manually). However, this approach is hard in terms of implementation, because it
requires extending the compiler. Also, this approach does not support combining
multiple extensions into one.
